---
title: "R_Notes_StudAss24"
author: "Sharon"
date: '`r Sys.Date()`'
output: html_document
---

```{r}
use_git()
```


# Basics

```{r working-directory, echo=F, eval=FALSE, results='hide', warning=F}
# what is the working directory?
getwd()

# if you need to adjust it:
setwd("/Users/sharonvandenburg/Documents/Werk & Loopbaan/Projecten/Stoepplantjes 2024/Data_R")

# set working dir as a variable
my.dir <- getwd()

# what files are in that working directory?
list.files(my.dir, recursive= TRUE)

# remove my.dir variable from environment
rm(my.dir)
```

```{r Packages, warning=F, results='hide', message=FALSE}
#layout file/export
library(knitr)
library(rmarkdown)
library(officer)
library(rvg)
library(xlsx)

# data
library(readxl)
library(dplyr)
library(plyr)
library(tidyverse)
library(vegan)
```

# Floron Data

## km-Grid cell 

```{r Import}
 km_Observation_data <- readRDS("~/Documents/Werk & Loopbaan/Projecten/Stoepplantjes 2024/Data_R/Laurens/Observation_data.rds")
 
 km_Visit_data <- readRDS("~/Documents/Werk & Loopbaan/Projecten/Stoepplantjes 2024/Data_R/Laurens/Visit_data.rds")
 
#also read "~/Documents/Werk & Loopbaan/Projecten/Stoepplantjes 2024/Data_R/Laurens/README_file.rds", zie wat hieronder staat
```

km_Observation_data:
- Site: 	site ID of visit, 1 x 1 km grid cells
- Visit:	visit ID
- Year:		year of visit, period: 2012 - 2019
- Date:		date of visit, period: 2012 - 2019
- **Species:	species ID, as used by FLORON in their database**

km_Visit_data:
- Site: 	site ID of visit, 1 x 1 km grid cells
- Visit:	visit ID
- Year:		year of visit, period: 2012 - 2019
- Date:		date of visit, period: 2012 - 2019
- Observer:	observer ID, the observer that uploaded observations to the database
- N_observers:	number of observers for a single list of observations
- Duration:	number of hours spent observing
- **N_species:	raw number of unique species recorded during the visit**

De data van een grid cell (bv. 49) staat dus per voorkomende plantsoort (bv nr 200) in de Obs file. Dit is samengevoegd tot species richness per visit (bv. 1920) in de Visit file. 

Alle losse observaties staan dus in de Obs file & alle samenvattingen per visit is de Visit file.

> Er wordt dus enkel naar Y/N voorkomen gekeken per soort + species richness

```{r delete-files-for-time-being}
rm(km_Observation_data, 
   km_Visit_data)
```


# Stoep Data

## Cleaning, editing

https://www.geeksforgeeks.org/data-cleaning-in-r/

Laadt data in hoe je normaal doet --> hernoem --> combine location data with occurrence data --> edit

```{r Import-and-merge}
#IMPORT DATA STOEP + BWNR
Import_dataStoep <- read_excel("~/Documents/Werk & Loopbaan/Projecten/Stoepplantjes 2024/Data_R/Export_stoep_R.xlsx", 
    sheet = "originele data") 

Import_dataStoep_loc <- read_excel("~/Documents/Werk & Loopbaan/Projecten/Stoepplantjes 2024/Data_R/adres_stoep20240704-203311.xlsx", 
    sheet = "adres_stoep20240704-203311") #1st version, 07/2024

#COMBINE DATASETS
library(dplyr)

# Select specific columns from Import_dataStoep_loc.test
edit.selected_columns <- select(Import_dataStoep_loc, Afstand, Lat, Lon, Straat, Huisnummer, Toevoeging, Postcode, Plaats)

# Combine with Import_dataStoep to create a new data set
edit.datastoep.combine <- cbind(edit.selected_columns, Import_dataStoep)

column_names <- names(edit.datastoep.combine)
print(column_names)
```

```{r Edit-1-Columns}
#rename columns
edit.datastoep.renamed <- edit.datastoep.combine %>%
  dplyr::rename(
    BAG_dist = Afstand,
    lat = Lat,
    lon = Lon,
    street = Straat,
    house_nr = Huisnummer,
    addition = Toevoeging,
    zip = Postcode,
    city = Plaats,
    session = sessie,
    sp_nr = soortnummer,
    date = datum,
    original_address = plaatsnaam,
    remark = opmerking,
    certainty_imagerec = zeker,
    original_kmcell = kmhok,
    submission_method = invoermethode,
    country = land,
    imagerec = image_recognition,
    original_x = x,
    original_y = y,
    province = provincie,
    dutch_name = nednaam,
    sci_name = wetnaam,
    family = familie,
    red_list = rl,
    rarity = zzz 
  )

# order columns
edit.datastoep.reordered <- edit.datastoep.renamed %>%
  select(id, session, date, created, # session 
         sp_nr, dutch_name,sci_name, family, red_list, rarity, # species 
         validated, imagerec, certainty_imagerec, # validation
         street, house_nr, addition, zip, city, province, country, lat, lon, # relevant location  
         original_x, original_y, original_address, BAG_dist, original_kmcell, # old location
         submission_method, email, naam, remark) #other
```

```{r Edit-2-Resolution}
# take out an error: Zoeterwoude is not copied into the new location data (city column)
edit.datastoep.reordered$city[edit.datastoep.reordered$original_address == "Zoeterwoude"] <- "Zoeterwoude"

# add column for highest resolution available (= most specific scale in original address)
# Create the new column based on the conditions
edit.datastoep.av_res <- edit.datastoep.reordered %>%
  mutate(
    available_resolution = case_when(
      # Condition for exactly 4 digits
      str_detect(original_address, "\\b\\d{4}\\b") ~ "zipcode",
      
      # Condition for a comma followed by a space and exactly two capital letters
      str_detect(original_address, ", [A-Z]{2}") ~ "city",
      
      # Condition for no comma and not exactly 4 digits
      !str_detect(original_address, ",") & !str_detect(original_address, "\\b\\d{4}\\b") ~ "city",
      
      # Default case
      TRUE ~ "street"
    )
  )

# add column for 4 letters of zip only
edit.datastoep.av_res$zip_stripped <- substr(edit.datastoep.av_res$zip, 1, 4)
```

```{r Take-out-weird-streetnames}
# Create the new column based on the conditions
edit.datastoep.av_res <- edit.datastoep.av_res %>%
  mutate(
    streetname_validation = case_when(
      # Condition for no comma and not exactly 4 digits
      str_detect(available_resolution, "street") & 
        str_detect(original_address, "^\\d+,") ~ "CHECK",
      
      # Default case
      TRUE ~ "OK"
    )
  )

check_data <- edit.datastoep.av_res %>%
  filter(streetname_validation == "CHECK")
print(check_data) # only one data point: 125993 (in city: Elburg)
```

> take out id 125993? --> heeft als adres '16'

```{r Check-BAG-distance}
# Create the new column based on the conditions
edit.datastoep.av_res <- edit.datastoep.av_res %>%
  mutate(
    BAG_dist_check = case_when(
      # Condition for more than 2 digits before the decimal point
      str_detect(BAG_dist, "^\\d{3,}\\.\\d*") ~ "CHECK",
      
      # Default case
      TRUE ~ "OK"
    )
  )

check_BAG_dist <- edit.datastoep.av_res %>%
  filter(BAG_dist_check == "CHECK") %>%
  select(id, street, zip, city, BAG_dist, original_address)
print(check_BAG_dist)

library(xlsx)
write.xlsx(check_BAG_dist, file="check_BAG_dist.xlsx", sheetName="high_BAG_dist")
```

> discuss with Maarten, BWNR

```{r Edit-3-Relevant-columns}
# only keep relevant colums
edit.datastoep.relevant <- edit.datastoep.av_res %>%
  select(id, session, date, # session 
         sp_nr, dutch_name,sci_name, family, red_list, rarity, # species 
         validated, imagerec, certainty_imagerec, # validation
         street, house_nr, addition, zip, zip_stripped, city, province, country, available_resolution, lat, lon # relevant location
  )
```

```{r Cleaning}
library(dplyr)
# clean data
clean.datastoep.validated <- edit.datastoep.relevant %>%
  filter(validated != 0) # exclude non-validated data (0=non validated)

clean.datastoep.NL <- clean.datastoep.validated %>%
  filter(country != 'BE') # exclude data BE

# rename clean data to simpler name
datastoep_complete <- clean.datastoep.NL #includes data of at least city resolution

# data set with only data with resolution of at least zip
datastoep_highresolution <- datastoep_complete %>%
  filter(available_resolution %in% c("zipcode", "street"))
```

```{r Clean-environment}
# Delete multiple datasets
rm(edit.selected_columns, 
   edit.datastoep.combine, 
   column_names,
   edit.datastoep.renamed,
   edit.datastoep.reordered,
   edit.datastoep.av_res,
   edit.datastoep.relevant,
   clean.datastoep.validated,
   clean.datastoep.NL,
   check_data,
   Import_dataStoep_loc, 
   Import_dataStoep
   )

# Check the remaining datasets
ls()
```

## Adding diveristy variables 

```{r Dataframes-&-Species-richness}
# ZIP 
library(tidyverse)
  # retrieve sci_name & location from main data set + make table with freq of indiv per sp per location
occurrence <- datastoep_highresolution[,c(6,16)] 
occurrence.table <- data.frame(table(occurrence)) 

library(dplyr)
  # make table with column per sp and occurrence data
sprich_zip <- occurrence.table  %>% 
  pivot_wider(
    names_from = sci_name,
    values_from = Freq
  )

library(plyr)
  # calculate sprich and put in a new dataset called diveristy
diversity_zip <- ddply(sprich_zip,~zip,function(x) {
   data.frame(sprich=sum(x[-1]>0))
 })

# ZIP STRIPPED 
library(tidyverse)
  # retrieve sci_name & location from main data set + make table with freq of indiv per sp per location
occurrence <- datastoep_highresolution[,c(6,17)] 
occurrence.table <- data.frame(table(occurrence)) 

library(dplyr)
  # make table with column per sp and occurrence data
sprich_zipstr <- occurrence.table  %>% 
  pivot_wider(
    names_from = sci_name,
    values_from = Freq
  )

library(plyr)
  # calculate sprich and put in a new dataset called diveristy
diversity_zipstripped <- ddply(sprich_zipstr,~zip_stripped,function(x) {
   data.frame(sprich=sum(x[-1]>0))
 })

# CITY 
library(tidyverse)
  # retrieve sci_name & location from main data set + make table with freq of indiv per sp per location
occurrence <- datastoep_complete[,c(6,18)] 
occurrence.table <- data.frame(table(occurrence)) 

library(dplyr)
  # make table with column per sp and occurrence data
sprich_city <- occurrence.table  %>% 
  pivot_wider(
    names_from = sci_name,
    values_from = Freq
  )

library(plyr)
  # calculate sprich and put in a new dataset called diveristy
diversity_city <- ddply(sprich_city,~city,function(x) {
   data.frame(sprich=sum(x[-1]>0))
 })
```

```{r Abundance}
# Zip 
ps <- table(datastoep_highresolution$sci_name, datastoep_highresolution$zip)
abundance <- colSums(ps)

    # Convert 'abundance' to a named vector with city names as names
abundance_named <- setNames(abundance, colnames(ps))
    # Merge 'diversity_city' with 'abundance_named'
diversity_zip <- diversity_zip %>%
  mutate(abundance = abundance_named[zip])

# Zip stripped
ps <- table(datastoep_highresolution$sci_name, datastoep_highresolution$zip_stripped)
abundance <- colSums(ps)

abundance_named <- setNames(abundance, colnames(ps))
diversity_zipstripped <- diversity_zipstripped %>%
  mutate(abundance = abundance_named[zip_stripped])

# City 
ps <- table(datastoep_complete$sci_name, datastoep_complete$city)
abundance <- colSums(ps)

abundance_named <- setNames(abundance, colnames(ps))
diversity_city <- diversity_city %>%
  mutate(abundance = abundance_named[city])
```

> voor volgende diveristy measures, zie Abundance voor code 

```{r Clean-environment-2}
# Delete multiple data sets
rm(
   ps,
   sprich_zip,
   sprich_zipstr,
   sprich_city,
   abundance,
   abundance_named,
   occurrence,
   occurrence.table,
   check_BAG_dist
   )

# Check the remaining data sets
ls()
```

## Location selection 

```{r #-locations-sufficient-data}
# zip
zip_30 <- sum(diversity_zip$abundance > 30) #21
zip_50 <- sum(diversity_zip$abundance > 50) #10
zip_100 <- sum(diversity_zip$abundance > 100) #3

# zip stripped
zipstr_30 <- sum(diversity_zipstripped$abundance > 30) #45
zipstr_50 <- sum(diversity_zipstripped$abundance > 50) #18
zipstr_100 <- sum(diversity_zipstripped$abundance > 100) #7

# city
city_30 <- sum(diversity_city$abundance > 30) #51
city_50 <- sum(diversity_city$abundance > 50) #27
city_100 <- sum(diversity_city$abundance > 100) #12
```

```{r table-locations-sufficient-data}
values <- c("zip_30", "zip_50", "zip_100", "zipstr_30", "zipstr_50", "zipstr_100", "city_30", "city_50", "city_100") 
sums <- c(zip_30, zip_50, zip_100, zipstr_30, zipstr_50, zipstr_100, city_30, city_50, city_100) 

# Extract thresholds and categories
thresholds <- sapply(strsplit(values, "_"), `[`, 2)
categories <- sapply(strsplit(values, "_"), `[`, 1)

# Map categories to their descriptions
category_descriptions <- ifelse(categories == "zip", "Zip",
                          ifelse(categories == "zipstr", "Zip stripped",
                                 ifelse(categories == "city", "City", NA)))

# Create a data frame
result_table <- data.frame(
   Scale = category_descriptions,
  "Min obs" = thresholds,
  "Num of loc" = sums
)

result_table$LessThan26 <- ifelse(result_table$Num.of.loc < 26, "Yes", "No")
result_table$LessThan21 <- ifelse(result_table$Num.of.loc < 21, "Yes", "No")
result_table$LessThan16 <- ifelse(result_table$Num.of.loc < 16, "Yes", "No")
result_table$LessThan11 <- ifelse(result_table$Num.of.loc < 11, "Yes", "No")

# Print the result table
print(result_table)
```

```{r export-locations-sufficient-data}
# >30
Locations_zip <- diversity_zip %>%
  filter(abundance > 30)
Locations_zipstr <- diversity_zipstripped %>%
  filter(abundance > 30)
Locations_city <- diversity_city %>%
  filter(abundance > 30)

# Export
library(xlsx)
write.xlsx(result_table, file="Location_selection_Resolution.xlsx", sheetName="#_per_resolution")
write.xlsx(Locations_zip,
           file="Location_selection_Resolution.xlsx",
           sheetName="zip_30", append=TRUE)
write.xlsx(Locations_zipstr,
           file="Location_selection_Resolution.xlsx",
           sheetName="zipstr_30", append=TRUE)
write.xlsx(Locations_city,
           file="Location_selection_Resolution.xlsx",
           sheetName="city_30", append=TRUE)

# >50
Locations_zip <- diversity_zip %>%
  filter(abundance > 50)
Locations_zipstr <- diversity_zipstripped %>%
  filter(abundance > 50)
Locations_city <- diversity_city %>%
  filter(abundance > 50)

# Add to export file
write.xlsx(Locations_zip,
           file="Location_selection_Resolution.xlsx",
           sheetName="zip_50", append=TRUE)
write.xlsx(Locations_zipstr,
           file="Location_selection_Resolution.xlsx",
           sheetName="zipstr_50", append=TRUE)
write.xlsx(Locations_city,
           file="Location_selection_Resolution.xlsx",
           sheetName="city_50", append=TRUE)

# >100
Locations_zip <- diversity_zip %>%
  filter(abundance > 100)
Locations_zipstr <- diversity_zipstripped %>%
  filter(abundance > 100)
Locations_city <- diversity_city %>%
  filter(abundance > 100)

# Add to export file
write.xlsx(Locations_zip,
           file="Location_selection_Resolution.xlsx",
           sheetName="zip_100", append=TRUE)
write.xlsx(Locations_zipstr,
           file="Location_selection_Resolution.xlsx",
           sheetName="zipstr_100", append=TRUE)
write.xlsx(Locations_city,
           file="Location_selection_Resolution.xlsx",
           sheetName="city_100", append=TRUE)
```

```{r clean-environment-3}
rm(
   values,
   city_100,
   city_50,
   city_30,
   zip_100,
   zip_50,
   zip_30,
   zipstr_100,
   zipstr_50,
   zipstr_30,
   category_descriptions,
   categories,
   sums,
   thresholds,
   result_table,
   Locations_city,
   Locations_zip,
   Locations_zipstr
)
```

















# Knit 

```{r Knitting, eval=F}
library(rmarkdown)
options(knitr.duplicate.label = "allow")
render('R_Notes_StudAss24_2.Rmd', 'all')
```

# Notes

## To add to diversity

A.S.A.P

* shannon? Niet logisch want abundance
* simpson? Niet logisch want abundance
* evenness? Niet logisch want abundance
* beta diversity! Hoe?

Later

* Functional div? Hoe? -> Hill numbers (hill_func_parti)

## Meeting LM 

```{r MEETING_LM, echo=F, eval=FALSE, results='hide', warning=F}
#Meeting LM
library(tidyverse)

datastoep.simple <- datastoep[,c(9,20)] # take sci_name & location from data set

data.stoep.table <- data.frame(table(datastoep.simple)) # make table with freq of indiv per sp per location

library(dplyr)

# make table with column per sp and occurrence data (zoals bij TUFF data)
data.stoep.table2 <- data.stoep.table  %>% 
  pivot_wider(
    names_from = wetnaam,
    values_from = Freq
  )
```

## To do

```{r Selection-locations}
# top ... locations --> under '#### Subsets top locations' in stage docx
```

```{r Graphs}
# Data per location within a (top 20) selection --> under '###### Graph type testing' in stage docx
```

## Other 

```{r Diveristy-datasets-19-8,echo=F, eval=FALSE, results='hide', warning=F }
# Zip:  d4, n2      - adding sp rich on zip level (thus high resolution needed)
unique_values <- unique(datastoep_highresolution$zip)
diversity_zip <- data.frame(zip = unique_values)

# Zip stripped:  d4 - adding sp rich on zip level (thus high resolution needed)
unique_values <- unique(datastoep_highresolution$zip_stripped)
diversity_zipstripped <- data.frame(zip_stripped = unique_values)

# City              - all data usable (thus datastoep_complete)
unique_values <- unique(datastoep_complete$city)
diversity_city <- data.frame(city = unique_values)


#{r Species-richness-OLD}

# see: https://rdrr.io/cran/vegan/man/diversity.html for explanation
library(vegan)

# Zip
sprich_select_zip <- table(datastoep_highresolution$sci_name,datastoep_highresolution$zip)
sprich_zip <- specnumber(sprich_select_zip,MARGIN=2)
sprich_zip <- as.data.frame(sprich_zip)

diversity_zip$sprich <- sprich_zip$sprich_zip # add to diveristy dataset

range(sprich_zip) # min, max value

# Zip stripped
sprich_select_zipstripped <- table(datastoep_highresolution$sci_name,datastoep_highresolution$zip_stripped)
sprich_zipstripped <- specnumber(sprich_select_zipstripped,MARGIN=2)
sprich_zipstripped <- as.data.frame(sprich_zipstripped)

diversity_zipstripped$sprich <- sprich_zipstripped$sprich_zipstripped # add to diveristy dataset

range(sprich_zipstripped) # min, max value

# See which would better to use > select on sp rich 
filtered_data_zip <- sprich_zip[sprich_zip$sprich_zip > 20, ]
filtered_data_zip_str <- sprich_zipstripped[sprich_zipstripped$sprich_zipstripped > 20, ] # meer data

    # > Waarom heeft Stripped meer data in de filter data dan niet stripped?! > logisch want door het samenvoegen van zips die dezelfde zip_stripped hebben zijn de getallen hoger ;)

# City 
sprich_select_city <- table(datastoep_complete$sci_name,datastoep_complete$city)
sprich_city <- specnumber(sprich_select_city,MARGIN=2)
sprich_city <- as.data.frame(sprich_city)

diversity_city$sprich <- sprich_city$sprich_city # add to diveristy dataset

#> de manier waarop ik de sprich variabel toevoeg aan de dataset klopt niet, dus ik heb het opniew geprobeerd (19/8)
```

```{r NOTES-add-to-dataset, echo=F, eval=FALSE, results='hide', warning=F}
print(ps)
print(abundance) # klopt gwn nog




# Convert 'abundance' to a named vector with city names as names
abundance_named <- setNames(abundance, colnames(ps))

# Merge 'diversity_city' with 'abundance_named'
diversity_city <- diversity_city %>%
  mutate(abundance = abundance_named[city])

# View the updated 'diversity_city' data frame
print(diversity_city)



abundance_named <- setNames(abundance, colnames(ps))
diversity_city <- diversity_city %>%
  mutate(abundance = abundance_named[city])
```

```{r Export-to-multiple-excel-sheets,echo=F, eval=FALSE, results='hide', warning=F}
library(openxlsx)
write.xlsx(check_BAG_dist, file="check_BAG_dist.xlsx", sheetName="high_BAG_dist", rowNames=TRUE)
write.xlsx(dataframe2, file="check_BAG_dist.xlsx", sheetName="sheet2", append=TRUE, rowNames=TRUE)
```

```{r table-location-selection,echo=F, eval=FALSE, results='hide', warning=F}
# Overview table 1
values <- c("zip_30", "zip_50", "zip_100", "zipstr_30", "zipstr_50", "zipstr_100", "city_30", "city_50", "city_100") 
sums <- c(zip_30, zip_50, zip_100, zipstr_30, zipstr_50, zipstr_100, city_30, city_50, city_100) 
  # Extract thresholds and categories
thresholds <- sapply(strsplit(values, "_"), `[`, 2)
categories <- sapply(strsplit(values, "_"), `[`, 1)
  # Map categories to their descriptions
category_descriptions <- ifelse(categories == "zip", "Zip",
                          ifelse(categories == "zipstr", "Zip stripped",
                                 ifelse(categories == "city", "City", NA)))
  # Create a data frame
result_table <- data.frame(
   Scale = category_descriptions,
  "Minimum observation" = thresholds,
  "Number of locations" = sums
)

print(result_table)


#Overview table 2
dataset_conditions <- c("> 30 observations", "> 50 observations", "> 100 observations")
resolution <- c("distance", "reaction", "distance")

# Calculate counts
counts <- c(sum(df$distance > 30), sum(df$reaction > 3), sum(df$distance > 40))

# Combine into a table
result_table <- data.frame(
  Condition = dataset_conditions,
  Count = counts,
  Column = columns
)

# Print the result table
print(result_table)
```

```{r sufficient-data, echo=F, eval=FALSE, results='hide', warning=F}
X <- 50
count <- sum(diversity_zip$abundance > X)
print(paste("Number of rows with values greater than", X, ":", count))
```

```{r Zoeterwoude-in-city-column-14/8, echo=F, eval=FALSE, results='hide', warning=F}
# Sample data frame
df <- data.frame(
  city = c("Amsterdam", "Rotterdam", "Zoetermeer"),
  original_address = c("Zoeterwoude", "Amsterdam", "Zoeterwoude")
)

print(df)

# Update the 'city' column where 'original_address' equals 'Zoeterwoude'
df$city[df$original_address == "Zoeterwoude"] <- "New City Name"

# Print updated data frame
print("Updated Data Frame:")
print(df)
```

```{r Meeting-LM, echo=F, eval=FALSE, results='hide', warning=F}
#Meeting LM
datastoep.simple <- datastoep[,c(9,20)] # haal wetnaam & locatie uit dataset

data.stoep.table <- data.frame(table(datastoep.simple)) # make table with freq of indiv per sp per location

library(dplyr)

# Combine Import data with location data 
datastoep.poging1 <- data.frame(
  Import_dataStoep[, c(1, 2, 5, 6, 7, 8, 9, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23)],  # Select columns from dataset1
  Import_dataStoep_loc.test[, c(4, 5, 6, 7, 8, 9, 10, 11)]   # Select columns from dataset2
)

# Select specific columns from dataset1
selected_columns <- select(Import_dataStoep_loc.test, Afstand, Lat, Lon, Straat, Huisnummer, Toevoeging, Postcode, Plaats)

# Combine with dataset2 to create a new dataset
datastoep.poging2 <- cbind(selected_columns, Import_dataStoep)

# make table with colum per sp and occurence data (zoals bij TUFF data)
data.stoep.table2 <- data.stoep.table  %>% 
  pivot_wider(
    names_from = wetnaam,
    values_from = Freq
  )
```

```{r old-code, echo=F, eval=FALSE, results='hide', warning=F}
# rename all colums
column_names <- names(edit.datastoep.combine)
print(column_names)

datastoep <- datastoep.combine %>%
  rename(
    new_name1 = Afstand,
    new_name2 = Lat,
    new_name3 = Lon,
    new_name1 = Straat,
    new_name2 = Huisnummer,
    new_name3 = Toevoeging,
    new_name1 = Postcode,
    new_name2 = Plaats,
    new_name3 = id,
    new_name1 = sessie,
    new_name2 = email,
    new_name3 = naam,
    new_name1 = soortnummer,
    date = datum,
    x = x,
    y = y,
    new_name2 = plaatsnaam,
    new_name3 = opmerking,
    new_name1 = created,
    new_name2 = validated,
    new_name3 = zeker,
    new_name1 = image_recognition,
    new_name2 = kmhok,
    new_name3 = invoermethode,
     new_name1 = land,
    new_name2 = provincie,
    new_name3 = nednaam,
    new_name1 = wetnaam,
    new_name2 = familie,
    new_name3 = rl,
    new_name1 = zzz 
  )

# Load dplyr package
library(dplyr)

# Example dataset
dataset.test <- data.frame(
  B = 1:3,
  A = 4:6,
  C = 7:9
)

# Reorder columns using dplyr
ordered_dataset.test <- dataset.test %>%
  select(A, B, C)

# View the reordered dataset
print(ordered_dataset.test)


# SPLIT COLUM
# Load tidyr package
library(tidyr)

# Example dataset
df <- data.frame(
  id = 1:3,
  text_column = c("A,1", "B,2", "C,3")
)

# Split the text_column into multiple columns
edit.datastoep.reordered2.seperate <- edit.datastoep.reordered2 %>%
  separate(original_address.dub, into = c("original_address.dub", "city?"), sep = ",")

# View the result
print(edit.datastoep.reordered2.seperate)
```

```{r available-resolution, echo=F, eval=FALSE, results='hide', warning=F}
# available resolution variable add

 - (1) add kolom with 'most specific scale available'
     - hiervoor heb ik het orginele adress nodig --> zit niet in datastoep_complete
     - hiervoor moet city uit het orginele adress gefilterd zijn --> doen in excel?
    
# trying to fix (1) --> zie hierboven 

# add colum that shows highest resolution scale available
```

```{r available-resolution-2, echo=F, eval=FALSE, results='hide', warning=F}
# Duplicate dataset (for now)
#edit.datastoep.reordered2 <- edit.datastoep.reordered 

# Duplicate Original location colum as column C
#edit.datastoep.reordered2$original_address <- edit.datastoep.reordered2$original_address.dub

# text to colum --> in hope to isolate city 


    # # Load tidyr package
    # library(tidyr)

    # Example dataset
    # df <- data.frame(
     id = 1:3,
     text_column = c("A,1", "B,2", "C,3")
    )

    # Split the text_column into multiple columns
    # edit.datastoep.reordered2.seperate <- edit.datastoep.reordered2 %>%
      separate(original_address.dub, into = c("original_address.dub", "city?"), sep = ",")

    # View the result
    # print(edit.datastoep.reordered2.seperate)

#USEFUL
  library(stringr)
 # Use stringr to split by the last comma
edit.datastoep.reordered.seperate <- edit.datastoep.reordered %>%
  mutate(
    # Create a new column with the part before the last comma
    before_last_comma = str_extract(original_address, "^[^,]+(,[^,]+)*"),
    # Create a new column with the part after the last comma
    after_last_comma = str_remove(original_address, "^[^,]+(,[^,]+)*,")
  ) 

! # Take out non cities
- filter 4 cijfers
- filter 2 Hoofdletters
- filter alle provincies

    edit.datastoep.reordered.seperate2 <- edit.datastoep.reordered.seperate %>%
      mutate(after_last_comma = str_c(" ", after_last_comma))

    # Modify the code to split by the last space (to take out spaces)
    edit.datastoep.reordered.separate3 <- edit.datastoep.reordered.seperate2 %>%
      mutate(
        # Create a new column with the part before the last space
       before_last_space = str_extract(original_address, "^[^ ]+( [^ ]+)*"),
        # Create a new column with the part after the last space
       after_last_space = str_remove(original_address, "^[^ ]+( [^ ]+)* ")
     )

# SEMI USEFUL DEZE # Create a new column excluding values that start with 4 digits or have exactly 2 capital letters
edit.datastoep.reordered.seperate.filter1 <- edit.datastoep.reordered.seperate %>%
  mutate(
    scale_filter1 = ifelse(
      !str_detect(after_last_comma, "^\\d{4}") & !str_detect(after_last_comma, "^[A-Z]{2}$"),
      after_last_comma,
      NA
    )
  )

#to do
- delete space at start
```

```{r available-resolution-3, echo=F, eval=FALSE, results='hide', warning=F}
edit.datastoep.reordered.separate.test <- edit.datastoep.reordered %>%
  mutate(
    # Create a new column with the part before the last comma
    before_last_comma = str_extract(original_address, "^(.*)(, [^,]*)?$"),
    
    # Create a new column with the part after the last comma, if present
    after_last_comma = str_remove(original_address, "^(.*)(, [^,]*)?$"),
    
    # If there is no comma, separate based on 4 digits, a space, and two capital letters
    before_last_comma = if_else(
      str_detect(original_address, ","),
      before_last_comma,
      str_extract(original_address, "^(\\d{4} [A-Z]{2})"),
    ),
    
    after_last_comma = if_else(
      str_detect(original_address, ","),
      after_last_comma,
      str_remove(original_address, "^(\\d{4} [A-Z]{2})")
    )
  )

# - ook nog DE eruithalen dan

# Updated code to handle both cases
edit.datastoep.reordered.separate.test <- edit.datastoep.reordered %>%
  mutate(
    # Split at the last comma
    before_last_comma = str_extract(original_address, "^(.*)(, [^,]*)?$"),
    after_last_comma = str_remove(original_address, "^(.*)(, [^,]*)?$"),
    
    # If there is no comma, split after 4 digits and a space
    before_digits_space = if_else(
      str_detect(original_address, ","),
      NA_character_, # NA if a comma is present
      str_extract(original_address, "^\\d{4} ")
    ),
    after_digits_space = if_else(
      str_detect(original_address, ","),
      NA_character_, # NA if a comma is present
      str_remove(original_address, "^\\d{4} ")
    )
  )

# Updated code to handle both cases
edit.datastoep.reordered.separate.test <- edit.datastoep.reordered %>%
  mutate(
    # Split at the last comma
    before_last_comma = str_extract(original_address, "^(.*)(, [^,]*)?$"),
    after_last_comma = str_remove(original_address, "^(.*)(, [^,]*)?$"),
    
    # If there is no comma, split after 4 digits and a space
    before_digits_space = if_else(
      str_detect(original_address, ","),
      NA_character_, # NA if a comma is present
      str_extract(original_address, "^\\d{4} ")
    ),
    after_digits_space = if_else(
      str_detect(original_address, ","),
      NA_character_, # NA if a comma is present
      str_remove(original_address, "^\\d{4} ")
    )
  )

```

```{r available-resolution-4, echo=F, eval=FALSE, results='hide', warning=F}
library(stringr)
 # Use stringr to split by the last comma
edit.datastoep.reordered.seperate <- edit.datastoep.reordered %>%
  mutate(
    # Create a new column with the part before the last comma
    before_last_comma = str_extract(original_address, "^[^,]+(,[^,]+)*"),
    # Create a new column with the part after the last comma
    after_last_comma = str_remove(original_address, "^[^,]+(,[^,]+)*,"),
    # Remove leading space from after_last_comma
    after_last_comma = str_trim(after_last_comma, side = "left"),
  ) 

# delete first 4 digits space 
# print into new column, except when 4 digits and space, or 4 digits space 2. capitals, or 2 capitals

```

```{r available-resolution-5, echo=F, eval=FALSE, results='hide', warning=F}
# create column that only contains a value when there is no zipcode or streetname 
# Create a new column based on the conditions
edit.datastoep.reordered.separate.poging2 <- edit.datastoep.reordered %>%
  mutate(
    # Copy values to the new column if there is no comma and not exactly 4 digits
    OnlyCityOrLower = if_else(
      str_detect(original_address, ",") | str_detect(original_address, "\\b\\d{4}\\b"),
      NA_character_, # NA if a comma is present or exactly 4 digits are found
      original_address
    )
  )
```

```{r available-resolution-6, echo=F, eval=FALSE, results='hide', warning=F}
# Create the new column based on the conditions
edit.datastoep.reordered.separate3 <- edit.datastoep.reordered %>%
  mutate(
    available_resolution = if_else(
      !str_detect(original_address, ",") & !str_detect(original_address, "\\b\\d{4}\\b"),
      "CityProvince",
      NA_character_ # Set to NA if the conditions are not met
    )
  )

# Create the new column based on the conditions
edit.datastoep.reordered.separate4 <- edit.datastoep.reordered %>%
  mutate(
    available_resolution = case_when(
      str_detect(original_address, "\\b\\d{4}\\b") ~ "zipcode",
      !str_detect(original_address, ",") & !str_detect(original_address, "\\b\\d{4}\\b") ~ "CityProvince",
      TRUE ~ "street" # Set to "street" for all other cases
    )
  )
```

```{r available-resolution-GOT-IT, echo=F, eval=FALSE, results='hide', warning=F}
# Create the new column based on the conditions
edit.datastoep.reordered.separate5 <- edit.datastoep.reordered %>%
  mutate(
    available_resolution = case_when(
      # Condition for exactly 4 digits
      str_detect(original_address, "\\b\\d{4}\\b") ~ "zipcode",
      
      # Condition for a comma followed by a space and exactly two capital letters
      str_detect(original_address, ", [A-Z]{2}") ~ "CityProvince",
      
      # Condition for no comma and not exactly 4 digits
      !str_detect(original_address, ",") & !str_detect(original_address, "\\b\\d{4}\\b") ~ "CityProvince",
      
      # Default case
      TRUE ~ "street"
    )
  )
```








# ...















